---
title: "Stage One Randomization Imbalance Simulations"
author: "John Sperger"
date: "9/17/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libs_and_scripts, include = FALSE}
library(tidyverse)
library(data.table)
library(fabricatr)
library(furrr)
if (str_detect(here::here(), "jsperger") == TRUE) plan(multisession, workers = 12) else plan(sequential)

source("./Code/minimization_algorithm.R")
source("./Code/generate_study_data.R")
source("./Code/analyze_imbalances.R")

```



## Minimization Metrics
Let $n_{A_{ij}}$ denote the number of patients that are on treatment A for the $j$-th level of prognostic factor $i$. Let $n_{A_{ijs}}$ denote the number of patients that are on treatment A for the $j$-th level of prognostic factor $i$ at site $s$. 
$$n_A = \sum_{j = 1}^{k_i}n_{A_{ij}}$$
Let $\mathcal{T}$ denote the set of treatments

#### Minimization Formulas

Marginal discrepancy between treatments $A$ and $B$

Including an overall term:
$$\text{MD} = w_0|n_A - n_B| + \sum_{i = 1}^K w_i\left(\sum_{j = 1}^{k_i} |n_{A_{ij}} - n_{B_{ij}}|\right)$$

Without an overall term:
$$\text{MD} = \sum_{i = 1}^K w_i\left(\sum_{j = 1}^{k_i} |n_{A_{ij}} - n_{B_{ij}}|\right)$$

#### Metrics
\begin{itemize}
\item Study-wide treatment imbalance: $$\max_{a, \, b \neq a \in \mathcal{T}}|n_a - n_b|$$

\item Study-wide prognostic factor: $$\max_{a, \, b \in \mathcal{T}}|n_{a_{ij}} - n_{b_{ij}}|$$

\item Within-site overall treatment imbalance: $$\max_{s} \max_{a, b \in \mathcal{T}}|n_{a_s} - n_{b_s}|$$

\item Within-site prognostic factor imbalance: $$\max_s \max_{a, \, b \in \mathcal{T}}|n_{a_{ijs}} - n_{b_{ijs}}|$$

\end{itemize}


### Accounting for contraindications



```{r sim_settings}
sim_reps <- 1e3
d_dim <- 4
covariate_dimensions <- 1:d_dim

#### Behavior flags
save_sim_data <- FALSE
sim_save_dir <- "./Results/"

use_saved_covar_data <- FALSE
saved_covar_path <- ""

#### Simulation Parameters
sim_seed <- 3990
# Seed from https://nclottery.com/Pick4 Daytime Draw Thursday, Sep 16

# Study parameters
n_subj <- 630
n_sites <- 10
k_arms <- 4

# Data generation parameters
coin_bias <- 2/3
bin_props = seq(from = .1, to = .9, length.out = length(covariate_dimensions))
bin_iccs = .05
contra_probs <- c(.55, .4, .025, .025) # None, dulox, pt, cbt

sim_run_settings <- list(sim_reps = sim_reps, 
                     covariate_dimensions = covariate_dimensions,
                     d_dim = d_dim,
                     n_subj = n_subj,
                     n_sites = n_sites,
                     k_arms = k_arms,
                     coin_bias = coin_bias,
                     bin_props = bin_props,
                     bin_iccs = bin_iccs,
                     contra_probs = contra_probs)
```

```{r sim_alt_settings}
# Define Alternative Settings

icc_settings <- c(.05, .15, .3)

dulox_ineligible_props <- seq(from = .3, to = .6, length.out = 3)

dulox_ineligible_settings <- map(dulox_ineligible_props, ~c(.x, 1 - (.x + .05), .025, .025))

settings_mixtures <- expand_grid(icc_settings, dulox_ineligible_settings)

icc_dulox_variant_settings <- map2(settings_mixtures$icc_settings, settings_mixtures$dulox_ineligible_settings, ~list_modify(sim_run_settings, bin_iccs = .x, contra_probs = .y))

```

```{r wrappers}
RunSimsFromSettingsList <- function(args.list){
  sim_data <-   map(1:args.list$sim_reps, ~GenerateStudyCovariateData(i = .,
                                               n.subj = args.list$n_subj,
                                               k.arms = args.list$k_arms, 
                                               n.sites = args.list$n_sites, 
                                               bin.props = args.list$bin_props, 
                                               bin.iccs = args.list$bin_iccs,
                                               contra.probs = args.list$contra_probs))
  
  return(sim_data)
}

ImbalanceWrapper <- function(df.list, d.dim){
  future_map_dfr(df.list, ~AnalyzeImbalances(indf = ., vars.to.check = paste0("X_", 1:d.dim)))
}

# Using the result of ImbalanceWrapper create a table with five number summaries of different imbalance measures
CreateImbalanceTable <- function(imbalances.df, method.name, d.dim){
  imbalances_long <- imbalances.df %>% pivot_longer(cols = ends_with("Imbalance"), 
                                           names_pattern = "(.*)Imbalance",
                                           names_to = "Comparison",
                                           values_to = "MaxImbalance")
  
  if("Covariates" %in% colnames(imbalances_long) == FALSE) imbalances_long$Covariates <- d.dim
  
  covar_summary <-imbalances_long %>% 
    mutate_at(., "Comparison", ~factor(., levels = c("MaxTreatment",
                                                     "MaxStudywideCov",
                                                     "MaxSiteTreatment",
                                                     "MaxSiteCov"),
                                       labels = c("Study-wide",
                                                  "Factor across study",
                                                  "Within-site",
                                                  "Factor within site"))) %>% group_by(Covariates, Comparison) %>% 
    group_modify(~ {
        .x %>%
            purrr::map_dfc(fivenum) %>%
            mutate(nms = c("Min", "Q1", "Median", "Q3", "Max"))
    }) %>% 
    pivot_wider(names_from = "nms", values_from = "MaxImbalance") %>% 
    arrange(Comparison, Covariates)
  
  covar_summary$Method <- method.name
  
  return(covar_summary)
}
```

```{r covariate_data_gen}
library(tictoc)
tic()
if(use_saved_covar_data == FALSE){
  covar_data_list <- future_map(seq_len(sim_reps), 
                            ~GenerateStudyCovariateData(i = ., 
                                               n.subj = n_subj,
                                               k.arms = k_arms, 
                                               n.sites = n_sites, 
                                               bin.props = bin_props, 
                                               bin.iccs = bin_iccs,
                                               contra.probs = contra_probs),
                            .options = furrr_options(seed = sim_seed))
  
  if(save_sim_data == TRUE){
    saveRDS(covar_data_list, file = paste0(sim_save_dir, lubridate::today(), "-sim_run.rds"))
  }
} else{
  covar_data_list <- read_rds(saved_covar_path)
}
toc()
```

```{r covaraite_data_gen_alt_assumptions}
library(tictoc)
tic()
  covar_data_variants_list <- future_map(icc_dulox_variant_settings, RunSimsFromSettingsList,
                            .options = furrr_options(seed = sim_seed))
  
toc()
```


```{r site_strat_with_contras, eval = FALSE}
tic()
# No stratification by site or contraindication, no overall imbalance term
contras_unstrat_no_overall <- map_dfr(covar_data_variants_list, 
                                  ~future_map_dfr(., 
                            ~GenerateAssignments(study.data = .,
                                                 n.sites = n_sites,
                                                 k.arms = k_arms, 
                                               stratify.by.site = FALSE),
                            .options = furrr_options(seed = sim_seed), .id = 'Replication'),
                            .id = 'Setting')
toc()

if(save_sim_data == TRUE) write_csv(x = contras_unstrat_no_overall,
                                    file = paste0(sim_save_dir, "unstrat_no_overall", lubridate::now(), ".csv"))

# No stratification by site or contraindication, includes an overall imbalance term
contras_unstrat_w_overall <- map_dfr(covar_data_variants_list, 
                                  ~future_map_dfr(., 
                            ~GenerateAssignments(study.data = .,
                                                 n.sites = n_sites,
                                                 k.arms = k_arms, 
                                               stratify.by.site = FALSE,
                                                use.overall.imbalance = TRUE),
                            .options = furrr_options(seed = sim_seed), .id = 'Replication'),
                            .id = 'Setting')

if(save_sim_data == TRUE) write_csv(x = contras_unstrat_w_overall,
                                    file = paste0(sim_save_dir, "unstrat_w_overall", lubridate::now(), ".csv"))

# Stratification by site but not contraindication, no overall imbalance term
contras_strat_by_site <- map_dfr(covar_data_variants_list, 
                                  ~future_map_dfr(., 
                            ~GenerateAssignments(study.data = .,
                                                 n.sites = n_sites,
                                                 k.arms = k_arms, 
                                               stratify.by.site = TRUE),
                            .options = furrr_options(seed = sim_seed), .id = 'Replication'),
                            .id = 'Setting')

if(save_sim_data == TRUE) write_csv(x = contras_strat_by_site,
                                    file = paste0(sim_save_dir, "strat_by_site", lubridate::now(), ".csv"))

# Stratification by site but not contraindication, no overall imbalance term
contras_strat_by_site_w_overall <- map_dfr(covar_data_variants_list, 
                                  ~future_map_dfr(., 
                            ~GenerateAssignments(study.data = .,
                                                 n.sites = n_sites,
                                                 k.arms = k_arms, 
                                               stratify.by.site = TRUE,
                                                use.overall.imbalance = TRUE),
                            .options = furrr_options(seed = sim_seed), .id = 'Replication'),
                            .id = 'Setting')

if(save_sim_data == TRUE) write_csv(x = contras_strat_by_site_w_overall,
                                    file = paste0(sim_save_dir, "strat_by_site_w_overall", lubridate::now(), ".csv"))
```

```{r site_strat_imbalances, cache=TRUE}
imbalances_contras_unstrat <- ImbalanceWrapper(contras_unstrat_no_overall, d_dim)
imbalances_contras_unstrat_overall_term <- ImbalanceWrapper(contras_unstrat_w_overall, d_dim)
imbalances_strat_by_site <- ImbalanceWrapper(contras_strat_by_site, d.dim = d_dim)
imbalances_strat_by_site_overall_term <- ImbalanceWrapper(contras_strat_by_site_w_overall, d.dim = d_dim)

imbalance_table_unstrat <- CreateImbalanceTable(imbalances_contras_unstrat, "Unstratified, no overall term", d_dim)
imbalance_table_overall_term <- CreateImbalanceTable(imbalances_contras_unstrat_overall_term, "Unstratified, overall term", d_dim)
imbalance_table_strat_by_site <- CreateImbalanceTable(imbalances_strat_by_site, "Stratified by site, no overall term", d_dim)
imbalance_table_strat_by_site_overall <- CreateImbalanceTable(imbalances_strat_by_site_overall_term, "Stratified by site, overall term", d_dim)
```

```{r dt_imbalance_summaries}
```